### 2016-11-29 paper/paper.tex
added some notes on method

#generate a paper tex file

#Method:
magSN = parallaxSN = 16 (including systematic parallax error in tgas)
minimum galactic latitude = 5 degrees
no NaNs in apass photometry, apass errors > 0
matched in Apass, wise, 2mass, rave
check that matching looks good with extreme colors
calculate distance modulus from sampling of gaia distances
correct for dust using Bayestar (Schlafly+Finkbeiner); minimal dust to matched stars
use absolute V band magnitude, B-V, g-r r-i colors

build KDTree of matched stars in Mv/color space
grab 2048 nearest neighbors
calculate chisq like distance metric between each chosen star and its neighbors
generate posterior belief of a chosen stars spectroscopic characteristics (ie Teff, logg, FeH)
  given the spectroscopic characteristics (and uncertainties) of its nearest neighbors in Mv/color space
.
.
.
.

#Discussion:
Apass HR Diagram looks very different from Rave HR Diagram, MS stars are incorrectly being labeled as giants