### 2016-12-28 paper/paper.tex
make paper to make

% This file is part of the stellarTwins project.
% Copyright 2016 the authors.

\documentclass[11pt,letterpaper]{article}

\begin{document}

\section*{Title here}

\noindent
Lauren Anderson (Flatiron) \& David W. Hogg (Flatiron)

\paragraph{Abstract:}
foo and bar

\section{Introduction}

Label transfer, The Cannon, etc.

\section{Method}

magSN = parallaxSN = 16 (including systematic parallax error in tgas)

minimum galactic latitude = 5 degrees

no NaNs in apass photometry, apass errors > 0

matched in Apass, wise, 2mass, rave

check that matching looks good with extreme colors

calculate distance modulus from sampling of gaia distances

correct for dust using Bayestar (Schlafly+Finkbeiner); minimal dust to matched stars

use absolute V band magnitude, B-V, g-r r-i colors

build KDTree of matched stars in Mv/color space

grab 2048 nearest neighbors

calculate chisq like distance metric between each chosen star and its neighbors

generate posterior belief of a chosen stars spectroscopic characteristics (ie Teff, logg, FeH)

given the spectroscopic characteristics (and uncertainties) of its nearest neighbors in Mv/color space

\section{Discussion}

Apass HR Diagram looks very different from Rave HR Diagram, MS stars are incorrectly being labeled as giants

\end{document}