### 2016-12-28 paper/paper.tex
typesetting tweaks

% This file is part of the stellarTwins project.
% Copyright 2016 the authors.
\documentclass[11pt,letterpaper]{article}

\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\Gaia}{\project{Gaia}}

\linespread{1.09}
\setlength{\parindent}{1.2\baselineskip}
\sloppy\sloppypar\raggedbottom\frenchspacing
\begin{document}

\section*{\raggedright%
A photometric method for obtaining spectroscopic parameters of stars}

\noindent
Lauren Anderson (Flatiron), David W. Hogg (Flatiron), others

\paragraph{Abstract:}
foo and bar

\section{Introduction}

Label transfer, The Cannon, etc.

\section{Method}

magSN = parallaxSN = 16 (including systematic parallax error in tgas)

minimum galactic latitude = 5 degrees

no NaNs in apass photometry, apass errors > 0

matched in Apass, wise, 2mass, rave

check that matching looks good with extreme colors

calculate distance modulus from sampling of gaia distances

correct for dust using Bayestar (Schlafly+Finkbeiner); minimal dust to matched stars

use absolute V band magnitude, B-V, g-r r-i colors

build KDTree of matched stars in Mv/color space

grab 2048 nearest neighbors

calculate chisq like distance metric between each chosen star and its neighbors

generate posterior belief of a chosen stars spectroscopic characteristics (ie Teff, logg, FeH)

given the spectroscopic characteristics (and uncertainties) of its nearest neighbors in Mv/color space

\section{Discussion}

Apass HR Diagram looks very different from Rave HR Diagram, MS stars are incorrectly being labeled as giants

\paragraph{Acknowledgements:}
This project was developed in part at the 2016 NYC Gaia Sprint, hosted
by the Center for Computational Astrophysics at the Simons Foundation
in New York City.

This work has made use of data from the European Space Agency (ESA)
mission \Gaia\ (http://www.cosmos.esa.int/gaia), processed by the \Gaia\ %
Data Processing and Analysis Consortium (\acronym{DPAC},
http://www.cosmos.esa.int/web/gaia/dpac/consortium). Funding for the
\acronym{DPAC} has been provided by national institutions, in particular the
institutions participating in the Gaia Multilateral Agreement.

\end{document}